---
title: "Capillary density analysis"
author:
- tabbassidaloii
output:
 md_document:
    variant: markdown_github
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

### Introduction
Here, we guide you through the step-by-step analysis process for quantifying capillary density. We use a human dataset to walk you through all the necessary step. To illustrate the steps, we will utilize a human dataset comprising muscle cryosections from 20 individuals. 

These sections were stained with CD31, CD105 (ENG), and laminin, and the imaging, image processing, and quantification procedures were conducted following the instructions outlined in the BioProtocol. 

It is important to note that all the steps, including laminin segmentation, were performed automatically, and for our purposes, we will consider the total area without the need for any additional filtration.

'Note:' The muscle abbreviations used were SEM (semitendinosus middle) and GAL(gastrocnemius lateralis) in all the analyses(scripts). However, they were only changed to STD and GL for the plots (shown in the publication).

```{r setup, include = FALSE}
# Empty the R environment
rm(list = ls())

# Set your working environment to the location where your current source file is saved into.
setwd(dirname(rstudioapi::getSourceEditorContext()$path))

# Check if libraries are already installed, otherwise install them
if(!"dplyr" %in% installed.packages()) install.packages("dplyr")
if(!"data.table" %in% installed.packages()) install.packages("data.table")
if(!"tidyr" %in% installed.packages()) install.packages("tidyr")
if(!"ggplot2" %in% installed.packages()) install.packages("ggplot2")
if(!"lmerTest" %in% installed.packages())install.packages("lmerTest")
if(!"knitr" %in% installed.packages()) install.packages("knitr")
if(!"reshape2" %in% installed.packages()) install.packages("reshape2")

# Load installed libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(data.table)
  library(tidyr)
  library(ggplot2)
  library(lmerTest)
  library(knitr)
})
```

#### Read the myofiber files and calculate the area
```{r myofibersTotalArea, echo = FALSE, message = FALSE, warning = FALSE, fig.show = FALSE, fig.height = 12, fig.width = 14}
# Read all the txt files
InputPath = "ROI/"
Files <- list.files(path = InputPath, pattern = ".txt", full.names = T)
DataTable <- lapply(Files, read.table)
names(DataTable) <- gsub(":.*", "", sapply(DataTable, function(x) x[1, 1]))

FiberData <- rbindlist(DataTable, idcol = TRUE) %>%
  select(.id, Label, Mean, Ch, Area) %>%
  unique() %>%
  mutate(Sample = gsub("-.*", "", .id)) %>%
  mutate(Replicate = gsub(".*_", "", .id)) %>%
  mutate(Individual = gsub("_.*", "", .id)) %>%
  mutate(Muscle = gsub("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  mutate(fiber = gsub(":.$", "", Label))

FiberData <- tidyr::pivot_wider(FiberData[,c("fiber", "Individual", "Muscle", "Replicate", "Ch", "Mean", "Area")], 
                                names_from = Ch, values_from = Mean)

No_Area_Fibers <- merge(FiberData %>%
  count(Muscle, Individual, Replicate) %>%
  rename(NoFiber = n), FiberData %>%
    group_by(Muscle, Individual, Replicate) %>%
    summarise(AreaFiber = sum(Area)))

FinalSamples <- No_Area_Fibers %>% 
  group_by(Muscle, Individual) %>%
  top_n(1, NoFiber) %>% #Select the replicate with the highest number of myofibers
  filter(Individual != "MD19") %>% #The GL sample had a few number of myofibers so samples from MD19 is excluded. 
  select(Muscle, Individual, Replicate)

rm(list = setdiff(ls(), c("DataTable", "col.muscle", "FiberData", "NoFibers", "No_Area_Fibers", "FinalSamples")))
```

#### Improt CD31 and CD105 quantification files, and calculate the %positive CD31 area
```{r positiveCD31Area, include = FALSE}
#Read all the txt files
InputPath = "Quantification/"
## CD31
Files <- list.files(path = InputPath, pattern = "CD31", full.names = T)

DataTable_CD31 <- lapply(Files, read.table)
names(DataTable_CD31) <- gsub(".*/|_AnalyzeParticles.txt", "", Files)
DataToFilt_CD31 <- rbindlist(DataTable_CD31, idcol = TRUE) %>% 
  select(- Label) %>%
  unique() %>%
  mutate(Sample = gsub("-.*", "", .id)) %>%
  mutate(Replicate = gsub("_.*", "", gsub(".*_s", "s", .id))) %>%
  mutate(Individual = gsub("_.*", "", .id)) %>%
  mutate(Muscle = gsub("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  rename(id_CD31 = .id, Mean_CD31 = Mean,
         StdDev_CD31 = StdDev, Mode_CD31 = Mode,
         Min_CD31 = Min, Max_CD31 = Max, Median_CD31 = Median,
         XM_CD31 = XM, YM_CD31 = YM)

## CD105
Files <- list.files(path = InputPath, pattern = "CD105", full.names = T)

DataTable_CD105 <- lapply(Files, read.table)
names(DataTable_CD105) <- gsub(".*/|_AnalyzeParticles.txt", "", Files)
DataToFilt_CD105 <- rbindlist(DataTable_CD105, idcol = TRUE) %>% 
  select(- Label) %>%
  unique() %>%
  mutate(Sample = gsub("-.*", "", .id)) %>%
  mutate(Replicate = gsub("_.*", "", gsub(".*_s", "s", .id))) %>%
  mutate(Individual = gsub("_.*", "", .id)) %>%
  mutate(Muscle = gsub("_.*", "", gsub("^[^_]*_", "", Sample))) %>%
  rename(id_CD105 = .id, Mean_CD105 = Mean,
         StdDev_CD105 = StdDev, Mode_CD105 = Mode,
         Min_CD105 = Min, Max_CD105 = Max,
         Median_CD105 = Median, XM_CD105 = XM, YM_CD105 = YM)

rm(list = setdiff(ls(), c("DataTable", "col.muscle", "FiberData", "NoFibers", "No_Area_Fibers", "FinalSamples", "DataTable_CD31", "DataTable_CD105")))

# Organize the tables
if(all(DataToFilt_CD105[ , c("Area", "Circ.", "AR", "Round", "Solidity", "Sample", "Replicate", "Individual", "Muscle")] == DataToFilt_CD31[, c("Area", "Circ.", "AR", "Round", "Solidity", "Sample","Replicate", "Individual", "Muscle")])) {
  DataToFilt <- cbind(DataToFilt_CD31, DataToFilt_CD105) %>% as.data.frame(as.is = T)
  for(col in c("Area", "Circ.", "AR", "Round", "Solidity", "Sample", "Replicate", "Individual", "Muscle")) {
    X = which(colnames(DataToFilt) == col)
    if(all(DataToFilt[, X[1]] == DataToFilt[, X[2]])) DataToFilt[, X[2]] = NULL
  }
}

# Calculate the area positive for CD31
AreaCD31 <- DataToFilt %>%
  group_by(Muscle, Individual, Replicate) %>%
  summarise(AreaCD31 = sum(Area))

# %CD31 area
ORDER <- c("GAL" = 2, "SEM" = 1)
col.muscle <- c("GL" = "#999999", "STM" = "#0066FF") #Abbreviations based on the paper
ggplot(merge(merge(No_Area_Fibers, AreaCD31), FinalSamples) %>%
         mutate(Order = ORDER[match(Muscle, names(ORDER))]) %>%
         mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")) %>%
         mutate(Pro_CD31Area = AreaCD31*100/AreaFiber),
       aes(x = reorder(Muscle, Order), y = Pro_CD31Area, fill = Muscle)) +
  geom_boxplot(position = position_dodge(0.8), alpha = 0.5) +
  stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
               geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "CD31 positive area(%)")  +
  theme(panel.grid = element_blank(), axis.text = element_text(size = 12), 
         axis.title.x = element_blank(), axis.title.y = element_text(size = 12), 
         axis.text.x.bottom = element_text(vjust = 0.5), axis.ticks = element_blank(),
         legend.position = "NULL")

flm <- as.formula("Pro_CD31Area ~ 0 + Muscle +(1|Individual)")
MM <- lmerTest::lmer(flm, data = merge(merge(No_Area_Fibers, AreaCD31), FinalSamples) %>% 
                                              mutate(Order = ORDER[match(Muscle, names(ORDER))]) %>%
                                              mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")) %>%
                                              mutate(Pro_CD31Area = AreaCD31*100/AreaFiber))
print(anova(MM)["Muscle",]["Pr(>F)"])*
print(fixef(MM))

```

#### Improt CD31 and CD105 quantification files, and calculate the %positive CD31 area
```{r capillaryDensity, include = FALSE}
# Filter
DataToFilt_CircArea <- DataToFilt %>%
  filter(Circ. > 0.5) %>%
  filter(3 < Area) %>%
  filter(Area < 51)

DataToFilt$Included <- ifelse(paste(DataToFilt$id_CD31, DataToFilt$Area, DataToFilt$Circ., DataToFilt$Max_CD31, DataToFilt$Min_CD31) %in% paste(DataToFilt_CircArea$id_CD31, DataToFilt_CircArea$Area, DataToFilt_CircArea$Circ., DataToFilt_CircArea$Max_CD31, DataToFilt_CircArea$Min_CD31), 1, 0)

colnames(DataToFilt) <- gsub("_CD31", "", colnames(DataToFilt))

# Keep capillaries positive for both CD31 and CD105
DataToFilt_CircArea_CD105 <- DataToFilt_CircArea %>%
  filter(Mean_CD105 != 0)

#Number of capillaries
NoCapillaries <- merge(DataToFilt_CircArea %>%
                     count(Muscle, Individual, Replicate) %>%
                     rename(NoCD31 = n),
                     DataToFilt_CircArea_CD105 %>%
                       count(Muscle, Individual, Replicate) %>%
                       rename(NoCD31_CD105 = n))

# Table includes the number of capillaries before and after filtering
AreaFibers_NoCapillaries <- merge(merge(NoCapillaries, No_Area_Fibers) %>%
  mutate(ratio =  NoCD31*100/AreaFiber,
          Filtered_ratio = NoCD31_CD105*100/AreaFiber), FinalSamples) 

# Capilary density
ggplot(AreaFibers_NoCapillaries %>%
          mutate(Order = ORDER[match(Muscle, names(ORDER))]) %>%
          mutate(Muscle = ifelse(Muscle == "GAL", "GL", "STM")),
        aes(x = reorder(Muscle, Order), y = Filtered_ratio, fill = Muscle)) + 
  geom_boxplot(position = position_dodge(0.8), alpha = 0.5) +
  stat_summary(fun.y = mean, color = "darkred", position = position_dodge(0.75),
                geom = "point", size = 3, show.legend = FALSE) +
  scale_fill_manual(values = col.muscle) +
  theme_bw() + labs(fill = element_blank(), y = "Capillary density")  +
  theme(panel.grid = element_blank(), axis.text = element_text(size = 12), 
         axis.title.x = element_blank(), axis.title.y = element_text(size = 12), 
         axis.text.x.bottom = element_text(vjust = 0.5), axis.ticks = element_blank(),
         legend.position = "NULL")

flm <- as.formula("Filtered_ratio ~ 0 + Muscle +(1|Individual)")
MM <- lmerTest::lmer(flm,
                      data = AreaFibers_NoCapillaries)
print(anova(MM)["Muscle",]["Pr(>F)"])
print(fixef(MM))
```

### Print session info
```{r, warning = FALSE, message = FALSE}
sessionInfo()
```
